### 2020-06-23 chapters/sota.tex
SOTA Sofia feedback applied

To find outliers, Extra-N has to do a linear search through the list of points and check the neighbor counts. Hence, for outlier detection, this method shows \textit{O(n)} complexity with \textit{n} as the number of points. Additionally, for each new data point \textit{p\textsubscript{i}}, Extra-N has to perform a range query search centered on it to find its neighbors and increment their lifetime neighbor counts by one. The run-time complexity of a range query search is \textit{O(log(n))} \cite{range-query-log} with \textit{n} as the number of data points.

@@

\section{Sliding Window Aggregations with Probabilistic Data Structures} \label{sec:sliding-pds}
In Section \ref{sec:pds} we presented a set of probabilistic data structures constant in time and space. These structures aggregate an endless stream of data. However, when working with sliding windows there is the need not only to insert new events but to evict old ones. Probabilistic data structures aggregate all incoming data but have no way of expiring old elements from the aggregation state.

@@

The work of Chabchoub et al. adapts the original HLL aggregator from Flajolet et al. to work under a sliding window framework  \cite{Chabchoub-Sliding-HLL}.
The main idea behind this implementation is to replace every HLL counter with a \textit{"List of Future Possible Maxima (LFPM)"}.

@@

Since each bucket now holds a LFPM instead of a single integer, the procedure for bucket update is different. Upon insertion of an element \textit{k}, a packet <t\textsubscript{k}, p\textsubscript{k}> is created with t\textsubscript{k} as the timestamp of arrival of the new element \textit{k} and p\textsubscript{k} as the position of the leftmost 1-bit position of \textit{k}'s binary representation. Given the new packet <t\textsubscript{k}, p\textsubscript{k}>, the authors propose the following LFPM update method, that checks for and evicts old packets at each insertion:

@@

The authors claim that the proposed sliding version remains as accurate as of the original HLL aggregator. In the context of their work, each element in the set is denoted as a "flow" and they state that \textit{"In practice, with an additional memory of only 35kB, a standard error of about 3\% can be achieved for a data stream of several million flows"}.

@@

The approach proposed by Xu uses a smaller amount of memory that does not change at run time \cite{Xu-hll-sliding-drv}. The author proposes a special counter: a \textit{"Distance Recorder Vector (DRV)"}.

@@

Initially, all records in the DRV will be set to be infinity or a large number.

\subsubsection*{Element Insertion}
When an element is inserted, the same procedure as in the original HLL is followed: the element is hashed and \textit{p} is computed from the hashed binary representation. Denote the bit-pattern observable of the new element as \textit{p\textsubscript{new}}. For the DRV of that bucket, increment the value of all records by one. Next, set position \textit{p\textsubscript{new}} in the DRV to 0 (DRV[\textit{p\textsubscript{new}}] = 0).

\subsubsection*{Sliding Window Cardinality Query} 
Element eviction is not performed. Instead, the query algorithm takes into account that some records in some of the DVR's might not be valid anymore. To estimate the sliding window cardinality is to find the value \textit{p}.

To find the value \textit{p} for a bucket, scan the DRV of that bucket backward. Start at the last element --- positioned at \textit{$\lceil ln(n/m) \rceil$} --- and scan towards the head of the vector. For each record, check if it is active in the current window. To do so, check if its value is less than the size of the window. If it is, the record is active. Otherwise, the record is inactive.

If the record is inactive continue scanning backward. If the record is active, then the position or index currently being scanned is the value of \textit{p} for that bucket. If all records are inactive, the bit-pattern observable \textit{p} is 0.

@@

Both inserting an element and estimating sliding window cardinality require a scan through \textit{$\lceil ln(n/m) \rceil$} records. This gives us a time complexity of \textit{O(\textit{ln(n)})}, given the number of buckets \textit{m} is constant.

>>>>

To find outliers, Extra-N has to do a linear search through the list of points and check the neighbor counts. Hence, for outlier detection, this method shows \textit{O(n)} complexity with \textit{n} as the number of points. Additionally, for each new data point \textit{p\textsubscript{i}}, Extra-N has to perform a range query search centered on it to find its neighbors and increment their lifetime neighbor counts by one. The runtime complexity of a range query search is \textit{O(log(n))} \cite{range-query-log} with \textit{n} as the number of data points.

@@

\section{Sliding Window Aggregations with Probabilistic Data Structures} \label{sec:sliding-pds}
In Section \ref{sec:pds} we presented a set of probabilistic data structures constant in time and space. These structures aggregate an endless stream of data. However, when working with sliding windows there is the need not only to insert new events but to evict old ones. The probabilistic data structures described until now can aggregate new incoming data but do not address the type of temporality needed for sliding windows --- \textit{i.e.} the ability to expire old elements from the aggregated state.

@@

The work of Chabchoub et al. adapts the original HLL aggregator from Flajolet et al. to work under a sliding window framework  \cite{Chabchoub-Sliding-HLL}.
The main idea behind this implementation is to replace every HLL counter with a \textit{List of Future Possible Maxima (LFPM)}.

@@

Since each bucket now holds a LFPM instead of a single integer, the procedure for bucket update is different. Upon insertion of an element \textit{k}, a packet $<t_k, p_k>$ is created with t\textsubscript{k} as the timestamp of arrival of the new element \textit{k} and p\textsubscript{k} as the position of the leftmost 1-bit position of \textit{k}'s binary representation. Given the new packet <t\textsubscript{k}, p\textsubscript{k}>, the authors propose the following LFPM update method, that checks for and evicts old packets at each insertion:

@@

The authors claim that the proposed sliding version remains as accurate as of the original HLL aggregator. With an additional memory of only 35kB it is possible to estimate the number of distinct elements in the set with a standard error of 3\% in the presence of up to a million distinct elements.

@@

The approach proposed by Xu uses a smaller amount of memory that does not change at run time \cite{Xu-hll-sliding-drv}. The author proposes a special counter: a \textit{Distance Recorder Vector (DRV)}.

@@

Initially, all records in the DRV will be set to be infinity or a very large number.

\subsubsection*{Element Insertion}
When an element is inserted, the same procedure as in the original HLL is followed: the element is hashed and \textit{p} is computed from the hashed binary representation. Let us denote the bit-pattern observable of the new element as \textit{p\textsubscript{new}}. For the DRV of that bucket, we increment the value of all records by one. Next, we set position \textit{p\textsubscript{new}} in the DRV to 0 --- \textit{i.e.} $\textrm{DRV}[p_{new}] = 0$.

\subsubsection*{Sliding Window Cardinality Query} 
This method does not perform element eviction. Instead, the query algorithm assumes that some records in DVR's entries might be expired. To estimate the sliding window cardinality is to find the value \textit{p}.

To find the value \textit{p} for a bucket, we scan the DRV of that bucket backward. For that, we start at the last element --- positioned at \textit{$\lceil ln(n/m) \rceil$} --- and scan towards the head of the vector. For each record, we check if it is active in the current window. To do so, we check if its value is less than the size of the window. If it is, the record is active. Otherwise, the record is inactive.

If the record is inactive we continue scanning backward. If the record is active, then the position or index currently being scanned is the value of \textit{p} for that bucket. If all records are inactive, the bit-pattern observable \textit{p} is 0.

@@

Both inserting an element and estimating sliding window cardinality require a scan through \textit{$\lceil ln(n/m) \rceil$} records. This gives us a time complexity of \textit{O(\textit{ln(n)})}, if the number of buckets \textit{m} is constant.
